---
title: "Hands-on Exercise 3: Geographical Segmentation with Spatially Constrained Clustering Techniques"
editor: visual
format: html
execute:
  warning: false
---

# Overview

This is a hands-on exercise based on [Chapter 5](https://r4gdsa.netlify.app/chap05.html "R4GDSA: Geographical Segmentation with Spatially Constrained Clustering Techniques") of [R for Geospatial Data Science and Analytics](https://r4gdsa.netlify.app/) by Dr. Kam Tin Seong and is a requirement under the class ISS624: Geospatial Analytics and Applications.

## Objectives

The objective of this hands-on exercise is to learn how to delineate homogeneous regions using geographically referenced multivariate data. There are two major analysis, namely:

-   hierarchical cluster analysis; and

spatially constrained cluster analysis.

By the end of this hands-on exercise, you will able:

-   to convert GIS polygon data into R's simple feature data.frame by using appropriate functions of **sf** package of R;

-   to convert simple feature data.frame into R's SpatialPolygonDataFrame object by using appropriate **sf** of package of R;

-   to perform custer analysis by using *hclust()* of Base R;

-   to perform spatially constrained cluster analysis using *skater()* of Base R; and

-   to visualise the analysis output by using **ggplot2** and **tmap** package.

## The Analytical Question

In geobusiness and spatial policy, it is a common practice to delineate the market or planning area into homogeneous regions by using multivariate data. In this hands-on exercise, we are interested to delineate [Shan State](https://en.wikipedia.org/wiki/Shan_State), [Myanmar](https://en.wikipedia.org/wiki/Myanmar) into homogeneous regions by using multiple Information and Communication technology (ICT) measures, namely: Radio, Television, Land line phone, Mobile phone, Computer, and Internet at home.

## The Datasets

-   Myanmar Township Boundary Data (i.e. *myanmar_township_boundaries*) : This is a GIS data in ESRI shapefile format. It consists of township boundary information of Myanmar. The spatial data are captured in polygon features. Under GIS Resources \> MIMU Geospatial Data.

-   *Shan-ICT.csv*: This is an extract of [**The 2014 Myanmar Population and Housing Census Myanmar**](https://myanmar.unfpa.org/en/publications/2014-population-and-housing-census-myanmar-data-sheet) at the township level.

Both data sets are download from [Myanmar Information Management Unit (MIMU)](http://themimu.info/)

# Getting Started

## The Required R Packages

The code chunk below installs and loads the different required packages for this exercise using `p_load()`:

```{r}
pacman::p_load(rgdal, spdep, tmap, sf, ggpubr, cluster, factoextra, NbClust, heatmaply, corrplot, psych, tidyverse)
```

::: {.callout-tip icon="false"}
## 🎮 LEVEL UP!

**NEW LIBRARIES UNLOCKED**!

-   `rgdal`

-   **`corrplot`**

-   **`ggpubr`**

-   **`heatmaply`**

-   **`cluster`**

-   **`NbClust`**

-   **`factoextra`**

-   **`psych`**
:::

## Importing the Data

### Importing geospatial data into the R environment

The code chunk below uses `st_read()` to import the shapefile containing the administrative boundaries of Myanmar.

```{r}
shan_sf <- st_read(dsn = "data/geospatial", layer = "myanmar_township_boundaries") %>% filter(ST %in% c("Shan (East)", "Shan (North)", "Shan (South)"))
```

::: callout-note
## 💻 CODE REVIEW!

**WHAT DOES THE OPERATOR `%in%` DO?**

So in the context of the code chunk above, the records are filtered to only extract those with `'ST'` equals to either "Shan (East)", "Shan (North)", "Shan (South)".

In the context of our data, `'ST'` refers to the state, region or union territory, which are the first level administrative boundaries of Myanmar. We are focusing on the "Shan" state of Myanmar.
:::

```{r}
shan_sf
```

```{r}
unique(shan_sf$TS)
```

As shown above we have 55 features, and each feature represents a township in Myanmar since the `'TS'` variable is unique for all records. The terms "feature", "polygon", and "township" will be used interchangeably in this exercise.

### Importing aspatial data into the R environment

The code chunk below uses `read_csv` to import

```{r}
ict <- read_csv("data/aspatial/Shan-ICT.csv")
```

```{r}
summary(ict)
```

The summary above confirms that there are 55 townships.

### Deriving new variables using **`dplyr`** package

```{r}
ict_derived <- ict %>% mutate(`RADIO_PR` = `Radio`/`Total households`*1000) %>% 
  mutate(`TV_PR` = `Television`/`Total households`*1000) %>% 
  mutate(`LLPHONE_PR` = `Land line phone`/`Total households`*1000) %>% 
  mutate(`MPHONE_PR` = `Mobile phone`/`Total households`*1000) %>%
  mutate(`COMPUTER_PR` = `Computer`/`Total households`*1000) %>% 
  mutate(`INTERNET_PR` = `Internet at home`/`Total households`*1000)
```

```{r}
ict_derived <- ict_derived  %>%  
  rename(`DT_PCODE` =`District Pcode`,
         `DT`=`District Name`,
         `TS_PCODE`=`Township Pcode`, 
         `TS`=`Township Name`,
         `TT_HOUSEHOLDS`=`Total households`, 
         `RADIO`=`Radio`, 
         `TV`=`Television`, 
         `LLPHONE`=`Land line phone`, 
         `MPHONE`=`Mobile phone`,
         `COMPUTER`=`Computer`, 
         `INTERNET`=`Internet at home`) 
```

```{r}
summary(ict_derived)
```

The new variables we created two code chunks ago (`'RADIO_PR'`, `'TV_PR'`, `'LLPHONE_PR'`, `'MPHONE_PR'`, `'COMPUTER_PR'`, and `'INTERNET_PR'`) are now in our new dataframe 'ict_derived'.

# Exploratory Data Analysis

## Original `'RADIO'` Distribution

```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO`)) + 
geom_histogram(bins=20, 
               color="#704276", 
               fill="#e3879e")
```

```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO`)) + 
geom_boxplot(color="#704276", 
             fill="#e3879e")
```

::: {.callout-important icon="false"}
## 🌸 FIRST IMPRESSIONS!

Based on the histogram above, it seems that most townships have around 10,000 households with radios, but there are a few townships that have around 30,000 households with radios. The data on households with radios looks skewed to the right.\
The succeeding bloxplot shows that there are 2 townships that are minor outliers and 1 township is a major outlier affecting the observed skewness from the histogram.
:::

## Derived `'RADIO_PR'` Distribution

```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="#704276", 
                 fill="#e3879e")
```

```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO_PR`)) + 
geom_boxplot(color="#704276", 
             fill="#e3879e")
```
